SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-03-27 2017-Xudong-SMT/ActaInformatica.tex
new proof for Th.2

\paragraph{Matching.}
The \emph{Matching} method builds the OTs of a pNet node from those of
its subnets.
For each synchronisation vector and each possible
combination of behaviours of the subnets, as generated by the \emph
{Combining} method, it builds the corresponding open transition.
Here, we only detail the construction of the predicate.
Suppose that $sv = \left((a'_i)^{i\in I} (b'_j)^{j\in J}\rightarrow v'\right) \in SV$ is a synchronisation vector, $G_k$ is its guard, and
$C = ({ot}_i)^{i \in [1,n]} \in \overline{comb}$ is a tuple of open transitions such that, for each $i \in [1,n]$, either $ot_i = -$, or the result action of $ot_i$ is $a_i$. Suppose futher that $\mathit{hole} = (b_j)^{j \in J}$ is the hole behaviour and $v$ is a fresh variable, representing the result action of the OT under construction. Then we can  build the predicate:
%
\begin{multline*}
  \mathit{MkPred}(sv, C, \overline{\mathit{hole}}) = 
  \\
  (\forall i\in I, a_i=a'_i) \land
%  (\forall i \not\in I, ot_i = -) \land
  (\forall j\in J, b_j=b'_j) \land
  (v=v') \land G_k.
\end{multline*}

@@

\begin{proof}
  Let $pn=<...>$ be a pNet. The set of states 
  managed by the two algorithms both are subsets of the cartesian
  product of the pLTS states, as defined in \pageref{def-states}. In a
  similar way, the set of state variables are also equal. Comparing
  transitions requires a little more care, because within the
  MakeTransitions function, fresh variables are created everytime a
  synchronisation vector is used (this occurs within the $MkPred$ and
  $matching$ functions). However, given a global state, the call of
  $MakeTransitions$ will always returns exactly the same set of OTs, up to
  renaming of these fresh variables\footnote{I think this does not
    need any specific proof, given that it is the same function,
    called with the same arguments}.

  Let us define the following sets of OTs, as computed by the 2
  algorithms:
  \ERIC{not sure whether all these sets are useful, but it's good to
    have them precisely defined for the moment}
  \begin{itemize}
    \item $L1$ = original set of transitions by Algorithm 1
      (i.e. before line 11), 
    \item $L1-sat$ = original set of transitions by Algorithm 1, after
      satisfiability check (i.e. after line 11),
    \item $L1-reach$ = final set of OTs of Algorithm 1, as computed
      by line 12.
    \item $L2-sat$ = original set of transitions by Algorithm 2
      (i.e. before line 13), 
    \item $L2-reach$ = final set of OTs of Algorithm 2, as computed
          by line 13.
  \end{itemize}

  For each of these, define $States : OT^* -> State^*$ the function
  computing the set of target states of a set of OTs.

  We want to prove that $L1-reach = L2-reach$ (upto renaming of fresh
  variables) and $States(L1-reach) = States(L2-reach)$.

  From the structure of the main loops of the algorithms, we know that
  $L1-sat \subset L2-sat$ and subsequently $States(L1-sat) \subset       
  States(L2-sat)$.
  \qed
\end{proof}

    \medskip
  \noteInEM{I keep below the original proof draft from Xudong, though
    I'm not sure if parts of it can be reused}
\begin{proof}
%% Let $L$, $L1$, $L_{Smart}$ respectively mean the result open transition set of original algorithm before pruning, original algorithm after pruning and smart algorithm. Similarly, $State(L)$, $State(L1)$, $State(L_{Smart})$ mean the result global state set of them. According to the algorithm, there has $L_{Smart}\in L$ and $State(L_{Smart})\in State(L)$.

We will construct a mapping $f$: $State(L1-reach)\rightarrow
State(L2-reach)$.
\ERIC{I do not understand this, as states are identical. You would
  rather need mapping betwen transitions ?}
\begin{enumerate}
\setlength{\itemindent}{3.5em}
	\item[(Surjective)] If $\ostate{\tilde{s_0}}=Initial(p)$ then $f(\ostate{\tilde{s}})=\ostate{\tilde{t_0}}=Initial(p)$ must exists, the transitions started from them fulfill that
	\[\forall \ostate{\tilde{t_0'}}, \ostate{\tilde{t_0}}\xrightarrow{v_j}\ostate{\tilde{t_0'}}\in L_{Smart},\quad \exists\ostate{\tilde{s'}},\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}\in L.\]
	The state $\ostate{\tilde{t'}}\in State(L_{Smart})$ and as $Pred_j$ of $\ostate{\tilde{t}}\xrightarrow{v_j}\ostate{\tilde{t_0'}}$ is satisfiable, then if the $Pred_i$ of $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}$ is also satisfiable, the state $\ostate{\tilde{s_0'}}$ is reachable in the original open automaton s.t. $\ostate{\tilde{s_0'}}\in State(L1)$, we can construct the mapping
	\[f(\ostate{\tilde{s_0'}})=\ostate{\tilde{t_0'}}.\]
	If the $Pred_i$ of $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}$ is not satisfiable then two transitions are not matched and $\ostate{\tilde{s_0'}}\notin State(L1)$.

	Let $\textit{Explored}$, $\textit{Explored}_{Smart}$ respectively mean the explored states set of original algorithm and smart algorithm.

	Suppose that  $\ostate{\tilde{s}}\in \textit{Explored}$ there is $\ostate{\tilde{t}}\in \textit{Explored}_{Smart}$ where $f(\ostate{\tilde{s}})=\ostate{\tilde{t}}$, there exists a sequence of transitions $\ostate{\tilde{s_0}}...\xrightarrow{v_i}...\ostate{\tilde{s}}$ and $\ostate{\tilde{t_0}}...\xrightarrow{v_j}...\ostate{\tilde{t}}$.

	The transitions $\ostate{\tilde{s}}\xrightarrow{v}\ostate{\tilde{s'}}$ and $\ostate{\tilde{t}}\xrightarrow{v}\ostate{\tilde{t'}}$ have several cases:
	\begin{enumerate}
		\item[(1)] $\ostate{\tilde{t'}}\in \textit{Explored}_{Smart}$

		We can find that 
		 \[\exists\ostate{\tilde{s'}}\in \textit{Explored},\quad \ostate{\tilde{s}}\xrightarrow{v}\ostate{\tilde{s'}}\quad where\quad f(\ostate{\tilde{s'}})=\ostate{\tilde{t'}}.\]
		\item[(2)] $\ostate{\tilde{t'}}\notin \textit{Explored}_{Smart}$

		There must has 
		\[\exists\ostate{\tilde{s'}}\notin \textit{Explored},\quad \ostate{\tilde{s}}\xrightarrow{v'}\ostate{\tilde{s'}}.\]
		If the $Pred_i$ of $\ostate{\tilde{s}}\xrightarrow{v'}\ostate{\tilde{s'}}$ is satisfiable then there exists a sequence of transitions $\ostate{\tilde{s_0}}...\xrightarrow{v_i}...\ostate{\tilde{s}}\xrightarrow{v'}\ostate{\tilde{s'}}$. So the $\ostate{\tilde{s'}}$ is reachable in the original open automaton, we can construct the mapping
		\[f(\ostate{\tilde{s'}})=\ostate{\tilde{t'}}.\]
		If the $Pred_i$ of $\ostate{\tilde{s}}\xrightarrow{v'}\ostate{\tilde{s'}}$ is not satisfiable then two transitions are not matched and $\ostate{\tilde{s'}}\notin State(L1)$.
	\end{enumerate}	
	Overall, \[\forall \ostate{\tilde{t}}\in State(L_{Smart}),\quad \exists\ostate{\tilde{s}}\in State(L1), f(\ostate{\tilde{s}})=\ostate{\tilde{t}}.\]
	\item[(Injective)] If there is a pair of states $\ostate{\tilde{s}} and \ostate{\tilde{t}}$ constructs the mapping $f(\ostate{\tilde{s}})=\ostate{\tilde{t}}$. We can know that $\ostate{\tilde{s}}\in \textit{Explored},\ostate{\tilde{t}}\in \textit{Explored}_{Smart}$.

	As when there is $\ostate{\tilde{t'}}\in \textit{Explored}_{Smart}$, the state is must be $\ostate{\tilde{s'}}\in \textit{Explored}$ where $f(\ostate{\tilde{s'}})=\ostate{\tilde{t'}}$.
	So it is a one-to-one mapping.
\end{enumerate}

Then as the function $f$ is both surjective and injective, it is a bijective function. The theorem is proved.

\end{proof}

>>>>

\paragraph{Matching.}
The \emph{Matching} method builds the OTs of a pNet node from those of
its subnets.
For each synchronisation vector and each possible
combination of behaviours of the subnets, as generated by the \emph
{Combining} method, it builds the corresponding open transition.
Here, we only detail the construction of the predicate.
Suppose that $sv = \left((a'_i)^{i\in I} (b'_j)^{j\in J}\rightarrow v'\right) \in SV$ is a synchronisation vector, $G_k$ is its guard, and
$C = ({ot}_i)^{i \in [1,n]} \in \overline{comb}$ is a tuple of open transitions such that, for each $i \in [1,n]$, either $ot_i = -$, or the result action of $ot_i$ is $a_i$. Suppose futher that $\mathit{hole} = (b_j)^{j \in J}$ is the hole behaviour and $v$ is a fresh variable, representing the result action of the OT under construction. Then we can  build the predicate:
%
\begin{multline*}
  \mathit{MkPred}(sv, C, \overline{\mathit{hole}}) = 
  (\forall i\in I, a_i=a'_i) \land
%  (\forall i \not\in I, ot_i = -) \land
  (\forall j\in J, b_j=b'_j) \land
  (v=v') \land G_k.
\end{multline*}

@@

%\begin{proof}
%  Let $pn=<...>$ be a pNet. The set of states 
%  managed by the two algorithms both are subsets of the cartesian
%  product of the pLTS states, as defined in \pageref{def-states}. In a
%  similar way, the set of state variables are also equal. Comparing
%  transitions requires a little more care, because within the
%  MakeTransitions function, fresh variables are created everytime a
%  synchronisation vector is used (this occurs within the $MkPred$ and
%  $matching$ functions). However, given a global state, the call of
%  $MakeTransitions$ will always returns exactly the same set of OTs, up to
%  renaming of these fresh variables\footnote{I think this does not
%    need any specific proof, given that it is the same function,
%    called with the same arguments}.
%
%  Let us define the following sets of OTs, as computed by the 2
%  algorithms:
%  \ERIC{not sure whether all these sets are useful, but it's good to
%    have them precisely defined for the moment}
%  \begin{itemize}
%    \item $L1$ = original set of transitions by Algorithm 1
%      (i.e. before line 11), 
%    \item $L1-sat$ = original set of transitions by Algorithm 1, after
%      satisfiability check (i.e. after line 11),
%    \item $L1-reach$ = final set of OTs of Algorithm 1, as computed
%      by line 12.
%    \item $L2-sat$ = original set of transitions by Algorithm 2
%      (i.e. before line 13), 
%    \item $L2-reach$ = final set of OTs of Algorithm 2, as computed
%          by line 13.
%  \end{itemize}
%
%  For each of these, define $States : OT^* -> State^*$ the function
%  computing the set of target states of a set of OTs.
%
%  We want to prove that $L1-reach = L2-reach$ (upto renaming of fresh
%  variables) and $States(L1-reach) = States(L2-reach)$.
%
%  From the structure of the main loops of the algorithms, we know that
%  $L1-sat \subset L2-sat$ and subsequently $States(L1-sat) \subset       
%  States(L2-sat)$.
%  \qed
%\end{proof}
%  
%    \medskip
%  \noteInEM{I keep below the original proof draft from Xudong, though
%    I'm not sure if parts of it can be reused}
\begin{proof}
Let
 \begin{itemize}
    \item $L1$ = original set of transitions by Algorithm 1
      (i.e. before line 11), 
    \item $L1_{sat}$ = original set of transitions by Algorithm 1, after
      satisfiability check (i.e. after line 11),
    \item $L1_{reach}$ = final set of OTs of Algorithm 1, as computed
      by line 12.
    \item $L2_{sat}$ = original set of transitions by Algorithm 2
      (i.e. before line 13), 
    \item $L2_{reach}$ = final set of OTs of Algorithm 2, as computed
          by line 13.
\end{itemize}
When the pNets contains no assignment, there is no need for the last filtering step, so $L1_{sat}=L1_{reach}$ and $L2_{sat}=L2_{reach}$.

For each of these, define $States: <OT> \rightarrow <State>$ which is the function computing the set of states in a set of OTs. Between two transitions sets of the
algorithms, define $Rename:  L1_{reach} \rightarrow L2_{reach}$ which is the function renaming fresh variables.

According to the rules Tr1 and Tr2 in \pageref{def-states}, the global state $\ostate{\tilde{s}}=\ostate{\tilde{s_0...s_n}}$ is a composition of n pLTS states $s_{i}^{i\in [0,n]}$.

From the structure of the main loop of the algorithms, we can get $States(L2_{reach})\subseteq  States(L1)$, then
\[\forall \ostate{\tilde{u}}\in States(L2_{reach}), \exists\ostate{\tilde{t}}\in States(L1), \ostate{\tilde{t}}=\ostate{\tilde{a_0...a_n}}=\ostate{\tilde{u}}.\]

Suppose that  $\ostate{\tilde{t}}\in States(L1_{reach})$, it is reachable in the result automaton of Algorithm 1, so there exists a sequence of transitions $\ostate{\tilde{t_0}}...\xrightarrow{v_i}...\ostate{\tilde{t}}$ where $Pred_i$ are all satisfiable. And there exists $\ostate{\tilde{u}}\in States(L2_{reach})$ where $\ostate{\tilde{t}}=\ostate{\tilde{u}}$.

Similarly, from the structure of the algorithms, 
\[\forall \ostate{\tilde{u}}\in States(L2_{reach}), \exists\ostate{\tilde{t}}\in States(L1), \ostate{\tilde{t}}=\ostate{\tilde{u}}\]
\[s.t.\quad OT_t = \ostate{\tilde{t}}\xrightarrow{v}\ostate{\tilde{t'}}\in L1, OT_u = \ostate{\tilde{u}}\xrightarrow{v}\ostate{\tilde{u'}}\in L2_{reach}, Rename(OT_t) = OT_u.\]
where $Pred_t$ and $Pred_u$ are satisfiable.

Then we can get a new longer sequence $\ostate{\tilde{t_0}}...\xrightarrow{v_i}...\ostate{\tilde{t'}}$ implying that $\ostate{\tilde{t'}}\in States(L1_{reach}),OT_t\in L1_{reach}$. So 
$States(L2_{reach})\subseteq  States(L1_{reach})$.

Moreover,
\[\forall OT_u\notin L2_{reach}, \exists OT_t\in L1, Rename(OT_t) = OT_u\]
where $Pred_t$ and $Pred_u$ are not satisfiable. Then according to the pruning function in Algorithm 1, $OT_t\notin L1_{reach}$.

As a result, we get that $L1_{reach}=L2_{reach}$, and subsequently, $States(L1_{reach})=States(L2_{reach})$.
%We will construct a mapping $f$: $State(L1-reach)\rightarrow
%State(L2-reach)$.
%\ERIC{I do not understand this, as states are identical. You would
%  rather need mapping betwen transitions ?}
%\begin{enumerate}
%\setlength{\itemindent}{3.5em}
%	\item[(Surjective)] If $\ostate{\tilde{s_0}}=Initial(p)$ then $f(\ostate{\tilde{s}})=\ostate{\tilde{t_0}}=Initial(p)$ must exists, the transitions started from them fulfill that
%	\[\forall \ostate{\tilde{t_0'}}, \ostate{\tilde{t_0}}\xrightarrow{v_j}\ostate{\tilde{t_0'}}\in L_{Smart},\quad \exists\ostate{\tilde{s'}},\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}\in L.\]
%	The state $\ostate{\tilde{t'}}\in State(L_{Smart})$ and as $Pred_j$ of $\ostate{\tilde{t}}\xrightarrow{v_j}\ostate{\tilde{t_0'}}$ is satisfiable, then if the $Pred_i$ of $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}$ is also satisfiable, the state $\ostate{\tilde{s_0'}}$ is reachable in the original open automaton s.t. $\ostate{\tilde{s_0'}}\in State(L1)$, we can construct the mapping
%	\[f(\ostate{\tilde{s_0'}})=\ostate{\tilde{t_0'}}.\]
%	If the $Pred_i$ of $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}$ is not satisfiable then two transitions are not matched and $\ostate{\tilde{s_0'}}\notin State(L1)$.
%	
%	Let $\textit{Explored}$, $\textit{Explored}_{Smart}$ respectively mean the explored states set of original algorithm and smart algorithm.
%	
%	Suppose that  $\ostate{\tilde{s}}\in \textit{Explored}$ there is $\ostate{\tilde{t}}\in \textit{Explored}_{Smart}$ where $f(\ostate{\tilde{s}})=\ostate{\tilde{t}}$, there exists a sequence of transitions $\ostate{\tilde{s_0}}...\xrightarrow{v_i}...\ostate{\tilde{s}}$ and $\ostate{\tilde{t_0}}...\xrightarrow{v_j}...\ostate{\tilde{t}}$. 
%	
%	The transitions $\ostate{\tilde{s}}\xrightarrow{v}\ostate{\tilde{s'}}$ and $\ostate{\tilde{t}}\xrightarrow{v}\ostate{\tilde{t'}}$ have several cases:
%	\begin{enumerate}
%		\item[(1)] $\ostate{\tilde{t'}}\in \textit{Explored}_{Smart}$
%		
%		We can find that 
%		 \[\exists\ostate{\tilde{s'}}\in \textit{Explored},\quad \ostate{\tilde{s}}\xrightarrow{v}\ostate{\tilde{s'}}\quad where\quad f(\ostate{\tilde{s'}})=\ostate{\tilde{t'}}.\]
%		\item[(2)] $\ostate{\tilde{t'}}\notin \textit{Explored}_{Smart}$
%		
%		There must has 
%		\[\exists\ostate{\tilde{s'}}\notin \textit{Explored},\quad \ostate{\tilde{s}}\xrightarrow{v'}\ostate{\tilde{s'}}.\]
%		If the $Pred_i$ of $\ostate{\tilde{s}}\xrightarrow{v'}\ostate{\tilde{s'}}$ is satisfiable then there exists a sequence of transitions $\ostate{\tilde{s_0}}...\xrightarrow{v_i}...\ostate{\tilde{s}}\xrightarrow{v'}\ostate{\tilde{s'}}$. So the $\ostate{\tilde{s'}}$ is reachable in the original open automaton, we can construct the mapping
%		\[f(\ostate{\tilde{s'}})=\ostate{\tilde{t'}}.\]
%		If the $Pred_i$ of $\ostate{\tilde{s}}\xrightarrow{v'}\ostate{\tilde{s'}}$ is not satisfiable then two transitions are not matched and $\ostate{\tilde{s'}}\notin State(L1)$.
%	\end{enumerate}	
%	Overall, \[\forall \ostate{\tilde{t}}\in State(L_{Smart}),\quad \exists\ostate{\tilde{s}}\in State(L1), f(\ostate{\tilde{s}})=\ostate{\tilde{t}}.\]
%	\item[(Injective)] If there is a pair of states $\ostate{\tilde{s}} and \ostate{\tilde{t}}$ constructs the mapping $f(\ostate{\tilde{s}})=\ostate{\tilde{t}}$. We can know that $\ostate{\tilde{s}}\in \textit{Explored},\ostate{\tilde{t}}\in \textit{Explored}_{Smart}$.
%	
%	As when there is $\ostate{\tilde{t'}}\in \textit{Explored}_{Smart}$, the state is must be $\ostate{\tilde{s'}}\in \textit{Explored}$ where $f(\ostate{\tilde{s'}})=\ostate{\tilde{t'}}$.
%	So it is a one-to-one mapping.
%\end{enumerate}
%
%Then as the function $f$ is both surjective and injective, it is a bijective function. The theorem is proved.
%
\end{proof}
